### 2018-11-12 manuscript/manuscript.tex
More work

\documentclass[review, a4paper, 10pt, times]{elsarticle}

\usepackage{lineno, hyperref}

\modulolinenumbers[5]

\journal{Agricultural and Forest Meteorology\tnoteref{mytitlenote}}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{A Multi-Satellite Investigation of the US Corn Belt in the Period 2010-2016 with a Focus on Solar-Induced Fluorescence}

%% Group authors per affiliation:
\author[uol,nceo]{Peter Somkuti\corref{cor1}\fnref{fn1}}
\ead{psomkuti@colostate.edu}
\author[uol,nceo]{Hartmut B\"osch}
\author[uoe,nceo]{Paul Palmer}
\author[uoe,nceo]{Liang Feng}
\author[uol,nceo]{Robert Parker}
\author[nceo]{Tristan Quaife}

\cortext[cor1]{Corresponding author, now at Colorado State University}
\fntext[fn1]{This is the specimen author footnote.}
\address[uol]{University of Leicester, Department of Physics of Astronomy, Leicester, UK}
\address[nceo]{National Centre for Earth Observation, University of Leicester, Leicester, UK}
\address[uoe]{University of Edinburgh, Edinburgh, UK}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\end{frontmatter}
\linenumbers

\section{Introduction}

Talk about Corn Belt, SIF, and other studies as well as their shortcomings. Mention especially the baseline problem.

\section{Materials \& Methods}
\subsection{Solar-Induced Fluorescence from GOSAT}
\label{sec:GOSAT_SIF}

The Greenhouse Gases Observing Satellite (GOSAT) is the first mission solely dedicated to measure atmospheric concentrations of carbon dioxide (CO$_2$) and method (CH$_4$). Launched in early 2009 and operated by the Japan Aerospace Exploration Agency (JAXA), GOSAT employs a Fourier-transform spectrometer (TANSO-FTS) to passively measure sunlight back-reflected from the surface of the Earth. These hyper-spectral measurements are then used together with appropriate physical models and inversion schemes to retrieve the atmospheric concentration of trace gases like CO$_2$, CH$_4$ and even H$_2$O.

Due to the high spectral resolution of TANSO-FTS in the O$_2$ A-band at wavelength of $760\:\mathrm{nm}$, several solar (Fraunhofer) lines can be resolved and utilised for the retrieval of SIF. We utilise the retrieval method described in \citet{Frankenberg2011}. This so-called \textit{physically-based} retrieval is derived from the Fraunhofer line discrimination method, first described in [REF]. Briefly summarised, the method relies on the fact that any additive radiation originating from the surface will reduce the fractional depth of Fraunhofer lines in the top-of-atmosphere (TOA) spectrum. It is therefore possible to decouple reflectance from the additive fluorescence contribution to the total TOA radiance. TANSO-FTS records radiances in the shortwave infrared wavelength range in two linear polarization states: P - parallel, and S - perpendicular to the plane of incidence (? or is it principal plane?).

To perform the GOSAT SIF retrievals we employ the University of Leicester Full-Physics algorithm \citep{Cogan2012}, called UoL-FP. It derives atmospheric and surface quantities from hyper-spectral measurements via its two components: a full-physics forward model, which accurately models the solar radiation entering and propagating through Earth's atmosphere, its interaction with the surface and finally the finite response at the detector on the space-based platform. The forward model calculations are then coupled with the second component of the algorithm: a non-linear inversion scheme that applies the Levenberg-Marquardt modification to the Gauss-Newton method, in order to modify the input parameters of the forward model (the so-called state vector) to match model calculation with real measurements. Inferring atmospheric and surface quantities as well as quantifying the uncertainty on each retrieved parameter is performed using a Bayesian optimal estimation (OE) framework according to \citet{Rodgers2000}.

We make use of the same processing pipeline which was previously utilized for the generation of high-quality atmospheric trace gas data sets \citep{Buchwitz2017,Trent2018} at the University of Leicester, using the same core retrieval algorithm. On a per-scene (measurement) basis, the surface pressure is informed by the ECMWF ERA-Interim model, and the surface albedo is estimated through the measured radiances themselves assuming a Lambertian surface model.

Following the retrieval strategy in \citet{Frankenberg2011}, we retrieve SIF at two micro-windows at $755\:\mathrm{nm}$ ($755.86 - 758.95$ nm) and $772\:\mathrm{nm}$ ($769.59 - 774.83$ nm), as well as the two polarizations, independently. For both retrieval windows, the SIF radiances are represented by a constant additive offset to the TOA radiances, along with a linear surface albedo (constant + slope), and instrument dispersion shift and stretch. As there are weak oxygen lines present in the $772\:\mathrm{nm}$ window, surface pressure is additionally retrieved in that micro-window. To exclude any scenes with large cloud contamination, we ran a cloud-screening algorithm beforehand. Only those GOSAT scenes are processed where the apparent surface pressure deviated by less than $100\:\mathrm{hPa}$ from the ERA-Interim prediction.

A crucial step in the SIF retrieval process is the bias correction. As established by \citet{Frankenberg2011}, small non-linearities in the analogue-to-digital conversion of the TANSO-FTS interferograms lead to the same fractional reduction of Fraunhofer lines as any SIF signal would do and thus result in unphysical SIF values. In order to rid the measurements of these instrument artefacts, the retrieved SIF values need to be corrected. To achieve the bias correction, we apply a slightly different method compared to \citet{Frankenberg2011}. First, we use the ESA CCI Land Cover map v2.0.7 \citep{Bontemps2013} to obtain the land cover classes for the surfaces seen by each GOSAT measurement (see Section~\ref{sec:collocation}). To identify footprints which can be assumed to be permanently free of vegetation, we consider only surfaces where at least $95\%$ of all land cover-class pixels within the footprint belong to the types: urban, desert, snow and ice, or any bare type. This approach has the advantage that not only highly reflective surfaces like ice shields or deserts are used for the correction process, but also darker surfaces such as mountainous regions.

The calibration curve is produced via the relationship between retrieved offset ($z_0$) and the mean radiance in the entire first GOSAT band ($\overline{I}_1$). Ideally, this curve would be flat, centered around zero, however the instrumental artefacts lead to a somewhat complicated shape, as shown in Figure XX [should I put a figure here? or is this too technical already..]. As this bias is of purely instrumental nature, it is not necessarily safe to assume that the bias remained the same over the long operational time of the instrument. Thus, we adopt three different strategies to correct for this instrumental bias. For the first one, we collect retrievals over non-fluorescing areas in seasonal aggregates (DJF, MAM, JJA, SON). For every season in every year the SIF retrievals are being corrected according to the calibration curve derived at that particular season-year combination. The second strategy is similar, the only difference being that retrievals for an entire year are collected. The main difference in the two methods is the aggregation period: shorter periods make the calibration curves noisy, however are more able to catch instrumental issues which may affect the bias. The third and most elaborate strategy takes the time-dependence more explicitly into account. First, the all retrievals over non-fluorescing areas are collected and binned two dimensionally: monthly bins for the time dimension, and $14$ bins in the radiance dimension. Through this regular 2D grid, a spline interpolation function is used to calculate the bias at any given point in time-radiance space.

\subsection{Spatially and Temporally Collocated Earth Observation Data}
\label{sec:collocation}

GOSAT/TANSO-FTS was designed to measure atmospheric quantities, and thus differs from imaging-type measurement concepts (e.g.~MODIS) that are tailored to the measurement of surface quantities. The high-resolution spectrometer points at one location at a time with an instantaneous field of view (IFOV) of ZYX$^\circ$, which corresponds roughly to a circular footprint with $10.5\:\mathrm{km}$ diameter at nadir viewing geometry. As a consequence, it is necessary to consider the specific sampling pattern for the data sets in the analysis in order to avoid sampling bias (see Section~\ref{sec:sampling_bias}).

For every GOSAT scene, we consider the full footprint as defined by the four extremal points of the footprint ellipse (as given in the L1B data). The following Earth observation datasets are then sampled:

\begin{itemize}
\item ESA CCI Land Cover v2.0.7 ($300\:\mathrm{m}$ resolution, $2010-2015$, for scenes after $2015$, the $2015$ map was used) \citep{Bontemps2013}
\item ESA CCI Soil Moisture v3.2 ($0.25^\circ$ resolution, $2010-2015$) \citep{Dorigo2017}
\item MODIS/Aqua MYD13C1 V006 Vegetation Indices ($0.05^\circ$ resolution, $2010-2016$) \citep{MYD13C1}
\item MODIS/Aqua MYD11C2 V006 Land Surface Temeprature ($0.05^\circ$ resolution, $2010-2016$) \citep{MYD11C2}
\item SPOT/VEGETATION Leaf Area Index V2.0.1 ($1\:\mathrm{km}$ resolution, $2010-2016$)
\item SPOT/VEGETATION Fractional Vegetation Cover V2.0.1 ($1\:\mathrm{km}$ resolution, $2009-2016$)
\item TRMM (3B43, multi-satellite) precipitation ($0.25^\circ$ resolution, $2010-2016$)
\end{itemize}

For the MODIS/Aqua, ESA CCI Land Cover and SPOT/VEGETATION datasets, the scene-specific footprint is fully taken into account. Only those pixels are aggregated to the scene-specific value, whose pixel area overlaps by at least $50\%$ with the GOSAT footprint. Soil moisture and TRMM precipitation data, however, are of lower resolution where sub-pixel aggregation is not required. For TRMM, the grid box is taken in which the GOSAT footprint lies, and for the soil moisture data (which is supplied on grid points rather than grid boxes [DOUBLE CHECK!!]), bi-linear interpolation of the closest four grid points is used.

In the time dimension, linear interpolation is used for all data sets except for land cover data (annual) and TRMM (monthly).

\subsection{Carbon Fluxes from Inversion of GOSAT CO2 Retrievals}
\label{sec:fluxes}
LIANG PLEASE HELP!

\subsection{Agricultural Statistics from USDA}
\label{sec:usda}

\subsection{GOSAT Sampling Bias}
\label{sec:sampling_bias}

As mentioned during the discussion on GOSAT SIF retrievals (Section~\ref{sec:GOSAT_SIF}), the FTS instrument, due to its pointing strategy/nature, causes potential sampling biases due to the pointing locations. While there is an inherent bias due to clouds, as GOSAT SIF retrievals can not be performed in the presence of thick clouds, the sampling strategy results in non-uniform selection of surfaces throughout the region of interest (ROI). Illustrated in Figure~\ref{fig:sampling}, the sampling density is shown for the entire time period from the beginning of $2010$ until the end of $2016$.

\begin{figure}[htbp]
\centering
\includegraphics{../coverage_map.png}
\caption{Land cover classes (left) and GOSAT sampling density shown for the region of interest ($104$W - $87.5W$, $37$N - $49$N). Yellow land cover pixels correspond to crop, orange represents grassland, green-colored pixels are various tree types, and red areas are urban regions. In the right picture, all GOSAT soundings for the entire time period $2010-2016$ are collected and visualized as a density map to illustrate the both sparse and irregular sampling pattern.}
\label{fig:sampling}
\end{figure}

The question arises, whether the sampling of the various surfaces of the chosen region of interest is representative of the entire area. To that end, we compare the most common land cover classes (Figure~\ref{fig:sampling} left) of the entire region of interest at the native resolution ($300\:\mathrm{m}$) of the ESA CCI Land Cover data set (excluding water bodies) to the sub-pixel aggregated data at the location of the GOSAT footprints. The result of the relative occurrence of the six most common types is shown in Figure~\ref{fig:lc_coverage}.

\begin{figure}[htbp]
\centering
\includegraphics{../lc_coverage.pdf}
\caption{default}
\label{fig:lc_coverage}
\end{figure}

Not surprising, more than $75\%$ of surfaces belong to either rainfed cropland or grassland types. The third most common surface type (deciduous tree cover) is mainly due to Ozark National Forest in Missouri, Superior National Forest as well as forests in northern Wisconsin. For the total period, GOSAT is well representing the overall composition of land cover classes of the region of interest. The largest differences are seen for flooded tree cover, where GOSAT is slightly over-repensenting ($+1.69\%$), and grassland, where GOSAT is under-representing ($-3.25\%$).

Repeating the same analysis on a monthly basis, a different picture than Figure~\ref{fig:lc_coverage} can arise, as GOSAT's sampling is not only non-uniform geographically, but also changed over time. In Figure~\ref{fig:lc_diff_monthly}, the differences between GOSAT-seen and total ROI land cover class coverage for monthly aggregates is shown for the three most prevalent classes. We observe that differences can change somewhat drastically, and for certain months even reach $10\%$. Months in which the differences are fairly large for cropland and grassland LC types are: $2013/04$, $2014/02$, $2014/03$, $2014/04$, $2014/11$, $2015/02$ and $2015/04$, $2016/08$. Due to the careful handling of the EO data set sampling, the difference in LC coverage should not have a significant impact on the further analysis and results. It is also important to note that while Figures~\ref{fig:lc_coverage} and \ref{fig:lc_diff_monthly} confirm that the LC classes are more or less well-represented, vegetation responses can vary geographically over the Corn Belt, even though they might share the same LC classification.

\begin{figure}[htbp]
\centering
\includegraphics{../lc_diff_monthly.pdf}
\caption{Jan 2015 has not been taken into account due to the low number of measurements in that month.}
\label{fig:lc_diff_monthly}
\end{figure}

\section{Results}

\subsection{Time Series}
\label{sec:SIF_TS}

The SIF retrievals are aggregated on a monthly basis, where the mean of all measurements inside the ROI for a given month are considered. As we have two different polarisations (P, S) and three different calibration procedures (seasonal, annual, spline-based), there are a total of six SIF time series. Figure~\ref{fig:sif_ts} shows this ensemble of SIF data, where the thick line is the median of all six time series, and the shaded area visualizes the minimum and maximum value of the given month to show the spread of the ROI-averaged SIF. It should be noted that the variability in the SIF time series stems mostly from the two different polarisations, rather than the calibration methodology. Months with less than $200$ measurements are dropped from both the visualization and the analysis, which affects only January $2015$, a month in which GOSAT was only partially operative due to instrumental issues.

\begin{figure}[htbp]
\centering
\includegraphics{../SIF_overview.pdf}
\caption{default}
\label{fig:sif_ts}
\end{figure}

The anomalies in Figure~\ref{fig:sif_ts} (bottom) are calculated as the difference between the current month and the median of all of the same months over the seven-year period. While the straight time series does not immediately reveal any striking features, the anomalies show several important ones. First, it is clear that over this relatively short period of seven years, the number of anomalous months within that period somewhat large: counting the number of months where the SIF signal exceeds the standard deviation of all anomalies (dotted green lines in Fig.~\ref{fig:sif_ts}, ), one arrives at $22$ (or $26\%$) anomalous months. We conclude that this short time period does not have a strong baseline, as every year has at least two anomalous months. [REFERENCE WOLF LIU AND OTHERS AND POINT OUT THAT THEY USED A SIMPLE BASELINE] It is therefore imperative to look at every anomaly within the context of other climatological and agricultural drivers.In chronological order, the big anomalies can be easily identified and related to known meteorological events.

Starting in $2010$, we see the effects of the rapid El Ni\~no-La Ni\~na transition, which brought excessive rain to the Corn Belt in the summer months (Fig.~\ref{fig:anomalies_0} bottom row) and drove plant growth.

In $2011$, we see a negative anomaly for spring, despite average precipitation and temperatures for that season. The reason for the negative anomaly is hinted at by the positive GRACE anomaly for $2011$ (Fig.~\ref{fig:anomalies_0} middle row). Large amounts of melting snow in the northern Corn Belt, as well as high precipitation in the south eastern corner (Missouri) have caused massive flooding of the Mississippi and Missouri rivers. As most parts of the ROI have not seen any increased rainfall, the anomaly plot in Figure~\ref{fig:anomalies_0} (bottom) does not show any significant outlier. The flooded areas, however, are visible as a positive anomaly in the GRACE data. The flooding has not only delayed planting progress, but also SOMETHING

The two anomalies in $2012$ are the most well-known and most-studied ones. In short, a warm winter and warm spring (see LST anomalies in Fig.~\ref{fig:anomalies_2} top) caused early and fast vegetation growth. In the following summer, however, anomalously low precipitation (Fig.~\ref{fig:anomalies_0} bottom) was responsible for drought conditions. As a consequence of the early spring onset due to warm temperatures, the plants used up soil water reserves quicker and depleted the soil of moisture, thus exacerbating the drought even further. This $2012$ drought was one of the most severe droughts in recorded history, with conditions comparable to the $1930$'s dustbowl era. The drought conditions are seen equally strongly in the soil moisture (Fig.~\ref{fig:anomalies_2} middle and bottom) anomalies [WHY IS PASSIVE NOT AS NICE HERE?], as well as the GRACE anomalies (Fig.~\ref{fig:anomalies_0} middle).

In $2013$ and $2014$, we again see a negative anomaly for spring and early summer. For these two years, the anomalies have similar causes. In the spring months, high precipitation and cool temperatures delayed the planting process - corn seed planting is generally done on warm and dry days - as well as caused slow maturing of crops. [POINT TO USDA GRAPHS].

The following two years exhibit positive SIF anomalies. In $2015$, near-ideal temperatures and favourable rainfall have led to good growing an maturing conditions for corn and soybeans, despite the rainfall distribution being suboptimal - the southern and eastern regions of the Corn Belt experienced more excessive rain. $2016$ proved to be an even better year for crop production than $2015$, and this is reflected in the SIF anomaly. After an intense El Ni\~no in $2015$, the quick transition to a weak La Ni\~na, and a fairly wet summer has provided ideal conditions for crops.

\begin{figure}[htbp]
\centering
\includegraphics{../anomalies_755_abs_0_median.pdf}
\caption{default}
\label{fig:anomalies_0}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{../anomalies_755_abs_2_median.pdf}
\caption{default}
\label{fig:anomalies_2}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{../anomalies_755_abs_1_median.pdf}
\caption{default}
\label{fig:anomalies_1}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{../anomalies_755_abs_3_median.pdf}
\caption{default}
\label{fig:anomalies_3}
\end{figure}

\section{Crop planting and yield}

The US Corn Belt is a very specific type of biome. As recently investigated by \citet{Alter2017}, it even generates its own climate. Despite the ROI being almost $1800 \cdot 10^9\;\mathrm{m}^2$ ($\approx 440$ million acres) in area, almost half of it is managed cropland. Meteorological conditions are thus not the only factor in determining the onset and productivity of one given season. It was pointed out in Section~\ref{sec:SIF_TS} that certain for years, the rainfall amount in springtime have caused delays in the planting of crops. To show that SIF indeed captures the delays in planting, we compare against the USDA crop progress reports. In these reports, the percentage of acreage planted is recorded based on farmer surveys. We translate these state-wide and weekly data for the whole ROI by multiplying with the total planted area for that given year and state in order to obtain an effective "planted area". We compare this number for every year against the integrated SIF radiances from Feburary onwards to June.

\begin{figure}[htbp]
\centering
\includegraphics{../corn_soy_planted_sif.pdf}
\caption{default}
\label{fig:corn_soy_planted_sif}
\end{figure}

The result for both corn and soybeans is shown in Figure~\ref{fig:corn_soy_planted_sif}. It shows that the time-integrated SIF radiances correspond very well to the crop planting progress of that given year at the first week of May (week 18). [TODO: If I change the months / weeks slightly - the result improves a bit; justification?]

\section*{}
\bibliography{mybibfile}

\end{document}